<<<<<<< HEAD
---
title: "HW 02 - Shot Charts"
author: "Sofia Guo"
date: "March 9, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = '../images/', eval = TRUE)
```

#5.1) Total Shots by Player
```{r}
library(dplyr)
library(magrittr)
library(readr)
shotsdata <- read.csv("../data/shots-data.csv", stringsAsFactors = FALSE)
#total number of shots by player, arranged in descending order
players <- group_by(shotsdata, name)
inc <- as_tibble(summarize(players,total = n()))
tot_shots <- arrange(inc, desc(total))
tot_shots
```
#5.2) Effective Shooting Percentage
```{r}
#effective shooting % by player
players <- group_by(shotsdata, name, shot_made_flag)
inc <- as_tibble(summarize(players,made = n()))
mshots <- filter(inc, shot_made_flag=='made shot')
decmshots <- arrange(mshots, desc(made))
decmshots$total <- tot_shots$total
woperc <- select(decmshots, c(1, 4, 3))
wperc <- mutate(woperc, perc_made = made/total*100)
arrange(wperc, desc(perc_made))

#2PT effective shooting % by player
players1 <- group_by(shotsdata, name, shot_type, shot_made_flag)
inc1 <- as_tibble(summarize(players1, made = n()))
mshots1 <- filter(inc1, shot_made_flag=='made shot', shot_type == '2PT Field Goal')
decmshots1 <- arrange(mshots1, desc(made))
decmshots1$total <- tot_shots$total[c(3,2,1,4,5)]
undecmshots1 <- ungroup(decmshots1)
woperc1 <- select(undecmshots1, c(1, 5, 4))
wperc1 <- mutate(woperc1, perc_made = made/total*100)
arrange(wperc1, desc(perc_made))

#3PT effective shooting % by player
players2 <- group_by(shotsdata, name, shot_type, shot_made_flag)
inc2 <- as_tibble(summarize(players2, made = n()))
mshots2 <- filter(inc2, shot_made_flag=='made shot', shot_type == '3PT Field Goal')
decmshots2 <- arrange(mshots2, desc(made))
decmshots2$total <- tot_shots$total
undecmshots2 <- ungroup(decmshots2)
woperc2 <- select(undecmshots2, c(1, 5, 4))
wperc2 <- mutate(woperc2, perc_made = made/total*100)
arrange(wperc2, desc(perc_made))
```
#6) Shooting Distance
##6.1) dplyr table
```{r}
#tibble with shot_distance and made_shot_prop

dist_prop <- shotsdata %>%
  group_by(shot_distance) %>%
  mutate(made_prop = ((shot_made_flag == 'made shot') / n())) %>%
  summarize(made_shot_prop = sum(made_prop))
dist_prop
```
##6.2) ggplot
```{r}
library(ggplot2)
ggplot(dist_prop, aes(shot_distance, made_shot_prop)) + geom_point()
```
Observations:
1. As the shot distance goes beyond 40, the proportion becomes extremely low (almost always zero). There is one outlier where the shot was made around 51 feet.
2. Preliminarily yes; judging from this data set, it is virtually impossible to make shots beyond 40 feet away from the basket.
3. My guesstimate is around 40 feet.
4. Distances below 5 have 50% or more chance of making it.

#7) Total number of shots by min of occurrence
```{r}
mins <- shotsdata %>%
  group_by(minute, name) %>%
  summarize(total_number_of_shots = n())

ggplot(mins, aes(minute, total_number_of_shots)) + annotate("rect", fill="gray", alpha = 0.5, xmin=seq(0, 50, 12), xmax=seq(12, 60, 12),ymin=0, ymax=50) + geom_point(color = 'dodgerblue1') + geom_path(color = 'dodgerblue1') + facet_wrap(~name) + theme_minimal() + scale_x_continuous(breaks = seq(0, 48, 12)) + ggtitle('Total number of shots (by minute of occurrence)') +scale_fill_gradient2(low = "red", high ="blue")


```

=======
---
title: "HW 02 - Shot Charts"
author: "Sofia Guo"
date: "March 9, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = '../images/', eval = TRUE)
```

#5.1) Total Shots by Player
```{r}
library(dplyr)
library(magrittr)
library(readr)
shotsdata <- read.csv("../data/shots-data.csv", stringsAsFactors = FALSE)
#total number of shots by player, arranged in descending order
players <- group_by(shotsdata, name)
inc <- as_tibble(summarize(players,total = n()))
tot_shots <- arrange(inc, desc(total))
tot_shots
```
#5.2) Effective Shooting Percentage
```{r}
#effective shooting % by player
players <- group_by(shotsdata, name, shot_made_flag)
inc <- as_tibble(summarize(players,made = n()))
mshots <- filter(inc, shot_made_flag=='made shot')
decmshots <- arrange(mshots, desc(made))
decmshots$total <- tot_shots$total
woperc <- select(decmshots, c(1, 4, 3))
wperc <- mutate(woperc, perc_made = made/total*100)
arrange(wperc, desc(perc_made))

#2PT effective shooting % by player
players1 <- group_by(shotsdata, name, shot_type, shot_made_flag)
inc1 <- as_tibble(summarize(players1, made = n()))
mshots1 <- filter(inc1, shot_made_flag=='made shot', shot_type == '2PT Field Goal')
decmshots1 <- arrange(mshots1, desc(made))
decmshots1$total <- tot_shots$total[c(3,2,1,4,5)]
undecmshots1 <- ungroup(decmshots1)
woperc1 <- select(undecmshots1, c(1, 5, 4))
wperc1 <- mutate(woperc1, perc_made = made/total*100)
arrange(wperc1, desc(perc_made))

#3PT effective shooting % by player
players2 <- group_by(shotsdata, name, shot_type, shot_made_flag)
inc2 <- as_tibble(summarize(players2, made = n()))
mshots2 <- filter(inc2, shot_made_flag=='made shot', shot_type == '3PT Field Goal')
decmshots2 <- arrange(mshots2, desc(made))
decmshots2$total <- tot_shots$total
undecmshots2 <- ungroup(decmshots2)
woperc2 <- select(undecmshots2, c(1, 5, 4))
wperc2 <- mutate(woperc2, perc_made = made/total*100)
arrange(wperc2, desc(perc_made))
```
#6) Shooting Distance
##6.1) dplyr table
```{r}
#tibble with shot_distance and made_shot_prop

dist_prop <- shotsdata %>%
  group_by(shot_distance) %>%
  mutate(made_prop = ((shot_made_flag == 'made shot') / n())) %>%
  summarize(made_shot_prop = sum(made_prop))
dist_prop
```
##6.2) ggplot
```{r}
library(ggplot2)
ggplot(dist_prop, aes(shot_distance, made_shot_prop)) + geom_point()
```
Observations:
1. As the shot distance goes beyond 40, the proportion becomes extremely low (almost always zero). There is one outlier where the shot was made around 51 feet.
2. Preliminarily yes; judging from this data set, it is virtually impossible to make shots beyond 40 feet away from the basket.
3. My guesstimate is around 40 feet.
4. Distances below 5 have 50% or more chance of making it.

#7) Total number of shots by min of occurrence
```{r}
mins <- shotsdata %>%
  group_by(minute, name) %>%
  summarize(total_number_of_shots = n())

ggplot(mins, aes(minute, total_number_of_shots)) + annotate("rect", fill="gray", alpha = 0.5, xmin=seq(0, 50, 12), xmax=seq(12, 60, 12),ymin=0, ymax=50) + geom_point(color = 'dodgerblue1') + geom_path(color = 'dodgerblue1') + facet_wrap(~name) + theme_minimal() + scale_x_continuous(breaks = seq(0, 48, 12)) + ggtitle('Total number of shots (by minute of occurrence)') +scale_fill_gradient2(low = "red", high ="blue")


```

>>>>>>> 98aebd82c21e4ca13d630d895fdffb40db07cde1
