---
title: "lab04-sofia-guo"
author: "Sofia"
date: "February 15, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Abalone Data Set Exploration

What is the charater delimiter?
Commas

Is there a row for column names?
No


Are there any missing values? If so, how are they codified?
No missing values

What is the data type of each column?
Sex is character; length, diameter, height, whole weight, shucked weight, viscera weight, shell weights are continuous. Rings are integers.

##Bash Command Exploration

Type of file: ASCII text, with CRLF line terminators

File Word Count:
4177 newline count, 4177 word count, 196050 byte count

To obtain only the number of lines, we use wc -l abalone.data

First 10 lines:
M,0.455,0.365,0.095,0.514,0.2245,0.101,0.15,15
M,0.35,0.265,0.09,0.2255,0.0995,0.0485,0.07,7
F,0.53,0.42,0.135,0.677,0.2565,0.1415,0.21,9
M,0.44,0.365,0.125,0.516,0.2155,0.114,0.155,10
I,0.33,0.255,0.08,0.205,0.0895,0.0395,0.055,7
I,0.425,0.3,0.095,0.3515,0.141,0.0775,0.12,8
F,0.53,0.415,0.15,0.7775,0.237,0.1415,0.33,20
F,0.545,0.425,0.125,0.768,0.294,0.1495,0.26,16
M,0.475,0.37,0.125,0.5095,0.2165,0.1125,0.165,9
F,0.55,0.44,0.15,0.8945,0.3145,0.151,0.32,19

To show only the first 5 lines, use head -5:
M,0.455,0.365,0.095,0.514,0.2245,0.101,0.15,15
M,0.35,0.265,0.09,0.2255,0.0995,0.0485,0.07,7
F,0.53,0.42,0.135,0.677,0.2565,0.1415,0.21,9
M,0.44,0.365,0.125,0.516,0.2155,0.114,0.155,10
I,0.33,0.255,0.08,0.205,0.0895,0.0395,0.055,7

Last 10 lines: 
M,0.5,0.38,0.125,0.577,0.269,0.1265,0.1535,9
F,0.515,0.4,0.125,0.615,0.2865,0.123,0.1765,8
M,0.52,0.385,0.165,0.791,0.375,0.18,0.1815,10
M,0.55,0.43,0.13,0.8395,0.3155,0.1955,0.2405,10
M,0.56,0.43,0.155,0.8675,0.4,0.172,0.229,8
F,0.565,0.45,0.165,0.887,0.37,0.239,0.249,11
M,0.59,0.44,0.135,0.966,0.439,0.2145,0.2605,10
M,0.6,0.475,0.205,1.176,0.5255,0.2875,0.308,9
F,0.625,0.485,0.15,1.0945,0.531,0.261,0.296,10
M,0.71,0.555,0.195,1.9485,0.9455,0.3765,0.495,12

Last 3 lines:
M,0.6,0.475,0.205,1.176,0.5255,0.2875,0.308,9
F,0.625,0.485,0.15,1.0945,0.531,0.261,0.296,10
M,0.71,0.555,0.195,1.9485,0.9455,0.3765,0.495,12

```{r}
abalone <- read.table("abalone.data", sep = ",")

head(abalone)
tail(abalone)
str(abalone, vec.len = 1)
```

#Creating names and data types for the data:

```{r}
column_names <- c(
  'sex',
  'length',
  'diameter',
  'height',
  'whole_weight',
  'shucked_weight',
  'viscera_weight',
  'shell_weight',
  'rings'
)

column_types <- c(
  'character',
  'real',
  'real',
  'real',
  'real',
  'real',
  'real',
  'real',
  'integer'   
)

abalone <- read.table(
  'abalone.data',
  col.names = column_names,
  colClasses = column_types,
  sep = ","
)

str(abalone, vec.len = 1)
```

#Assigning type and value to Abalone Table
```{r}
library(readr)

head(read.csv("abalone.data", col.names = column_names, colClasses = column_types))

ct <- 'cdddddddi'

head(read_csv("abalone.data", col_names = column_names, col_types = ct))


```

#Pittsburgh Bridges Data Set
Are there column names? 
Yes
What is the field separator? Commas
Are there any missing values?
Yes
What is the character for missing values (if any)?
?
What is the data type of each variable (i.e. column)?
River: n
Location: n
Erected: c,n (range of continuous numbers is given first and the possible values follow semicolon)
Purpose: n
Length: c,n
Lanes: c,n
Clear-G: n
T-or-D: n
Material: n
Span: n
Rel-L: n
Type: n

##Downloading the file:
download.file('http://archive.ics.uci.edu/ml/machine-learning-databases/bridges/bridges.data.version1', 'bridges.data')

```{r}
bridges1 <- read.table('bridges.data')
bridges2 <- read.csv('bridges.data')

#create vector of classes for first five columns
firstfive <- c(rep("character", 5), rep("NULL", 8))

bridges3 <- read.table('bridges.data', header = FALSE, sep = ",", colClasses = firstfive)
```
##Exploring bridges1
```{r}
str(bridges1)
summary(bridges1)
head(bridges1)
tail(bridges1)
dim(bridges1)
names(bridges1)
colnames(bridges1)
nrow(bridges1)
ncol(bridges1)
```

#Creating Data Frames
```{r}
#GSW Data Frames
Player <- c("Thompson", 
            "Curry", 
            "Green", 
            "Durant", 
            "Pachulia")
Position <- c('SG', 'PG', 'PF', 'SF', 'C')
Salary <- c(16663575, 12112359, 15330435, 26540100, 2898000)
Points <- c(1742, 1999, 776, 1555, 426)
PPG <- c(22.3, 25.3, 10.2, 25.1, 6.1)
Rookie <- c(rep('False', 5))
#data frame first
GSW.data <- data.frame(Player,
                       Position,
                       Salary,
                       Points,
                       PPG,
                       Rookie)
#list first
GSWlist.data <- list(Player,
                     Position,
                     Salary,
                     Points,
                     PPG,
                     Rookie)
GSW1 <- data.frame(GSWlist.data)
Position <- factor(Position)
class(Salary)
Points <- as.integer(Points)
class(Points)
class(PPG)
Rookie <- as.logical(Rookie)
#reassign GSW1 to reflect new data types
GSWlist.data <- list(Player,
                     Position,
                     Salary,
                     Points,
                     PPG,
                     Rookie)
GSW1 <- data.frame(GSWlist.data)

#Use cbind
bound <- cbind(Player, Position, Salary, Points, PPG, Rookie)

#convert to data frame
data.frame(bound)
```
#NBA Players
```{r}
dat <- read.csv('nba2017-players.csv', stringsAsFactors = FALSE)
```

Display the last 5 rows of the data.
```{r}
dat[446:441, ]
```


Display those rows associated to players having height less than 70 inches tall.
```{r}
heights70 <- dat[dat$'height' < 70, ] 
heights70
```


Of those players that are centers (position C), display their names and salaries.
```{r}
C <- dat[dat$'position' == 'C', ]
C[ ,c(1,9)]
```

Create a data frame durant with Kevin Durant's information (i.e. row).
```{r}
dat[dat$'player' == 'Kevin Durant', ]
```


Create a data frame ucla with the data of players from college UCLA ("University of California, Los Angeles").
```{r}
ucla <- dat[dat$college == "University of California, Los Angeles", ]
ucla
```


Create a data frame rookies with those players with 0 years of experience.
```{r}
rookies <- dat[dat$experience == 0, ]
rookies
```


Create a data frame rookie_centers with the data of Center rookie players.
```{r}
rookie_centers <- rookies[rookies$'position' == "C", ]
rookie_centers
```

Create a data frame top_players for players with more than 50 games and more than 100 minutes played.
```{r}
top_players <- dat[dat$minutes > 100 & dat$games > 50, ]
top_players
```


What's the largest height value?
```{r}
max(dat$height)
```


What's the minimum height value?
```{r}
min(dat$height)
```


What's the overall average height?
```{r}
mean(dat$height)
```


Who is the tallest player?
```{r}
tallest <- dat[dat$height == max(dat$height), ]
tallest[ ,1]
```


Who is the shortest player?
```{r}
shortest <- dat[dat$height == min(dat$height), ]
shortest[ ,1]
```

Which are the unique teams?
```{r}
unique(dat$team)
```

How many different teams?
```{r}
length(unique(dat$team))
```

Who is the oldest player?
```{r}
dat[dat$age == max(dat$age), ][ ,1]
```

What is the median salary of all players?
```{r}
median(dat$salary)
```


What is the median salary of the players with 10 years of experience or more?
```{r}
players10 <- dat[dat$experience <= 10, ]
median(players10$salary)
```

What is the median salary of Shooting Guards (SG) and Point Guards (PG)?
```{r}
SGPG <- dat[dat$position == 'SG' | dat$position == 'PG', ]
median(SGPG$salary)
```

What is the median salary of Power Forwards (PF), 29 years or older, and 74 inches tall or less?
```{r}
spec <- dat[dat$position == 'PF' | dat$age >= 29 | dat$height <= 74, ]
median(spec$salary)
```

How many players scored 4 points or less?
```{r}
nrow(dat[dat$points <= 4, ])
```

Who are those players who scored 4 points or less?
```{r}
dat[dat$points <= 4, 1]
```

Who is the player with 0 points?
```{r}
dat[dat$points == 0,1]
```

How many players are from "University of California, Berkeley"?
```{r}
nrow(dat[dat$college == "University of California, Berkeley", ])
```

Are there any players from "University of Notre Dame"? If so how many and who are they?
```{r}
ND <- dat[dat$college == "University of Notre Dame", 1]
ND
length(ND)
```

Are there any players with weight greater than 260 pounds? If so how many and who are they?
```{r}
weight260 <- dat[dat$weight > 260, 1]
weight260
length(weight260)
```

How many players did not attend a college in the US?
NBA Players must be drafted through US colleges

Who is the player with the maximum rate of points per minute?
```{r}
dat[,"rate"] <- dat[, 'points'] / dat[, 'minutes']

dat[dat$rate == max(dat$rate),1]

```

Who is the player with the maximum rate of three-points per minute?
```{r}
dat[,"rate3"] <- dat[, 'points3'] / dat[, 'minutes']

dat[dat$rate3 == max(dat$rate3),1]

```

Who is the player with the maximum rate of two-points per minute?
```{r}
dat[,"rate2"] <- dat[, 'points2'] / dat[, 'minutes']

dat[dat$rate2 == max(dat$rate2),1]

```

Who is the player with the maximum rate of one-points (free-throws) per minute?
```{r}
dat[,"rate1"] <- dat[, 'points1'] / dat[, 'minutes']

dat[dat$rate1 == max(dat$rate1),1]

```

Create a data frame gsw with the name, height, weight of Golden State Warriors (GSW)
```{r}
gsw <- dat[dat$team == 'GSW', c(1,4,5)]
gsw
```

Display the data in gsw sorted by height in increasing order (hint: see ?sort and ?order)
```{r}
gsw[order(gsw$height), ]
```

Display the data in gsw by weight in decreasing order (hint: see ?sort and ?order)
```{r}
gsw[order(gsw$weight, decreasing=TRUE), ]
```

Display the player name, team, and salary, of the top 5 highest-paid players (hint: see ?sort and ?order)
```{r}
salary_sorted <- dat[order(dat$salary, decreasing=TRUE), ]

salary_sorted[1:5, c('player', 'team', 'salary')]

```

Display the player name, team, and points3, of the top 10 three-point players (hint: see ?sort and ?order)
```{r}
points3_sorted <- dat[order(dat$points3, decreasing=TRUE), ]

salary_sorted[1:10, c('player', 'team', 'points3')]
```

Quick try: Using just bracket notation, try to create a data frame with two columns: the team name, and the team payroll (addition of all player salaries).
```{r}
unique_team <- dat[dat$team == unique(dat$team), ]
sum(unique_team$salary)
#it's pretty much impossible with just brackets because I need a way to group by team first.
```


Create a data frame with the average height, average weight, and average age, grouped by position
```{r}
aggregate(dat[ ,c('height', 'weight', 'age')], by = list(dat$position), FUN = mean)
```

Create a data frame with the average height, average weight, and average age, grouped by team
```{r}
aggregate(dat[, c('height', 'weight', 'age')], by = list(dat$team), FUN = mean)
```

Create a data frame with the average height, average weight, and average age, grouped by team and position.
```{r}
aggregate(dat[ ,c('height', 'weight', 'age')], by = list(dat$position,dat$team), FUN = mean)
```

Difficult: Create a data frame with the minimum salary, median salary, mean salary, and maximum salary, grouped by team and position.
```{r}

x <- dat[ ,'salary']
aggregate(x, by = list(dat$position, dat$team), FUN = function(x) c(mi = min(x), md = median(x), mn = mean(x), mx = max(x) ) )

```

